Supporting Information : Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-08-27 Ben_Manuscripts/structure_paper/supporting_information.tex
Reorganizing supporting info

@@

  \vspace{1em}
  We parameterized monomers according to the following procedure:
  \begin{enumerate}
	\item \textit{Create monomer structure file with connectivity} : We
	drew atomistic structures using MarvinSketch
	17.13~\cite{chemaxon_marvinsketch_2017} with all hydrogen atoms drawn out
	explicitly. We optimized the 3D geometry of the structure using the 'Clean in
	3D' function of MarvinSketch.  We saved the structure as a .mol file, then
	converted it to .pdb format using Open Babel 2.4.1
	\cite{oboyle_open_2011,noauthor_open_nodate}. 
	\item \textit{Assign GAFF atomtypes using \texttt{antechamber}} : Using
	the .pdb structure file as input, we ran
	\texttt{antechamber}~\cite{wang_automatic_2006} using the AM1-BCC charge model.
	The net charge on the monomer is input as -1 since the sodium ion is kept as 
	a separate residue. We use \texttt{LEaP}~\cite{case_ambertools16_2016} and the
	output of \texttt{antechamber} to create Amber topology files. A detailed
	tutorial can be accessed elsewhere \cite{walker_antechamber_nodate}.
	\item \textit{Create GROMACS topologies from Amber output} : The output
	of \texttt{LEaP} is a .inpcrd and a .prmtop file which are Amber topology
	files. Using acpype.py \cite{sousa_da_silva_acpype_2012}, we converted the
	\texttt{LEaP} output into GROMACS .gro and .top files. 
	\item \textit{Perform a simulated annealing procedure on the monomer} :
	We created a cubic box around the monomer using the GROMACS command \texttt{gmx
	editconf}. The monomer was centered in the box with edges of the
	box spaced at least 3 nm from the monomer on all sides. We ran an energy minimization
	on the system with the steepest descent algorithm. Next we performed an NVT
	simulated annealing procedure. We linearly decreased the temperature of the
	system from 1000 K to 50 K over the course of 10 ns. We randomly chose a monomer
	configuration from the last 10 \% of the trajectory. 
	\item \textit{Reassign charges with \texttt{molcharge}}: With the monomer
	configuration taken from the annealed trajectory, we reassigned charges using
	\texttt{molcharge} with the am1bccsym method in order to ensure charges 
	are symmetric. This condition is not guaranteed with \texttt{antechamber}. 
	The charges in the GROMACS topology file (.top) were replaced with the 
	new charges calculated by \texttt{molcharge}. 
	\item \textit{Anneal again to get final structure} : We repeated the
	same simulated annealing procedure using the monomer topology with
	\texttt{molcharge} charges. A random monomer configuration was pulled from the
	last 10 \% of the trajectory and was used to build all assemblies reported
	(Figure~\ref{fig:monomer}).
  \end{enumerate}

  \begin{figure}[!htb]
	\centering
        \includegraphics[width=0.9\textwidth]{monomer.png}
	\caption{Atomistic representation of the monomer Na-GA3C11. White atoms
		represent hydrogen, cyan atoms represent carbon, red atoms represent oxygen and
		the blue atom is sodium.}\label{fig:monomer}
  \end{figure}

  \begin{figure}[!htb]
	\centering
        \includegraphics[width=0.9\textwidth]{monomer_color_coded.png}
	\caption{The groups used for radial distribution calculations. Red atoms are in the
		tails group. Green atoms are in the head group region. The blue atom is sodium. 
		}\label{fig:monomer_color_coded}
  \end{figure}

  \begin{figure}[!htb]
        \centering
        \begin{subfigure}[b]{0.32\textwidth}
                \centering
                \includegraphics[width=\textwidth]{sandwiched.png}
                \caption{}\label{fig:sandwiched}
        \end{subfigure}
        \begin{subfigure}[b]{0.32\textwidth}
                \centering
                \includegraphics[width=\textwidth]{PD.png}
                \caption{}\label{fig:pd}
        \end{subfigure}
        \begin{subfigure}[b]{0.32\textwidth}
                \centering
                \includegraphics[width=\textwidth]{Tshaped.png}
                \caption{}\label{fig:tshaped}
        \end{subfigure}
        \vskip\baselineskip
        \begin{subfigure}[b]{0.475\textwidth}
                \centering
                \includegraphics[width=\textwidth]{sandwichedlayers.png}
                \caption{}\label{fig:sandwichedlayers}
        \end{subfigure}
        \begin{subfigure}[b]{0.475\textwidth}
                \centering
                \includegraphics[width=\textwidth]{offsetlayers.png}
                \caption{}\label{fig:offsetlayers}
        \end{subfigure}
        \caption{(a) Sandwiched benzene dimers stack 3.8 \angstrom~apart. (b) Parallel-Displaced benzene dimers stack
        3.4 \angstrom~vertically and 1.6 \angstrom~horizontally apart. (c) T-shaped benzene dimers stack 5.0 \angstrom~apart.
        (d) Two monomer layers stacked in the sandwiched configuration (e) Two monomer layers stacked in the parallel-displaced
        configuration }\label{fig:stacking}
  \end{figure}

  \begin{figure}[!htb]
	\centering
	\includegraphics[width=0.8\textwidth]{pore_radius_illustration.png}
	\caption{When creating an initial configuration, the pore radius is defined based
	on the distance of the carbonyl carbon from the pore's central axis}
	\label{fig:pore_radius_illustration}
  \end{figure}

  We chose to build all systems using 20 monomers per column. We made this choice
  by considering a number of factors that directly affect the quality of our results. 
  Clearly, choosing a small number of layers will cause the system to suffer from 
  finite size effects. Even in our 20 monomer-per-column system, axial correlations
  between stacked monomers persist throughout the entire correlation function (see
  Figure ~\ref{M-fig:correlation}). The correlation function of the system which we
  built with 40 monomers per column continues to oscillate for its full length
  (Figure~\ref{fig:tall_correlation_full}). If we continue adding monomers, the 
  correlation function will eventually fully decay but it is not worth the 
  computational expense. Regardless, the correlation lengths calculated from 20
  monomer per column simulations are in sufficient agreement with experiment.

  \begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{tall_correlation_full.png}
  \caption{Oscillations persist throughout the length of the correlation 
  function for the sandwiched system in the ordered basin built with 40 monomers
  per column. The full system is ca. 17 nm tall so the full correlation function
  covers half this length since periodicity forces the remaining length to be its
  mirror image.}\label{fig:tall_correlation_full}
  \end{figure}

  We may be able to simulate systems using less than 20 monomers per column, but chose
  to use 20 since it gives fairly good resolution when simulating X-ray diffraction 
  patterns. The size of the Fourier space bin in each dimension is determined by 
  $dfrac{2\pi}{L}$ where L is the length of a given box vector. The z-box vector 
  for 20 monomer-per-column systems is $\approx$ 85 \AA which means we will see 
  a reciprocal space resolution of 0.074 \AA$^{-1}$ in the z direction. Further decreasing
  the resolution would increase the noise in the simulated patterns and make it more
  difficult to measure the intensity of reflections of interest.

  \section{Initial Configuration Dependence}\label{section:initial_config_dependence}
%  \begingroup
%	\fontsize{14pt}{12pt}\selectfont
%	\textbf{Initial Configuration Dependence}
%  \endgroup

  \vspace{1em}
  We addressed any major dependence on initial configuration in the main text.
  There we showed that systems are stable when made with 4, 5, 6, 7, and 8
  monomers per layer. We also showed that systems are stable when monomer head
  groups are oriented in the parallel displaced and sandwiched configurations
  (see Figure~\ref{fig:stacking}). Our model is most consistent with experiment
  when built with 5 monomers per layer, with layers stacked 3.7 \AA~apart in 
  the parallel displaced configuration.

  There are three other parameters whose choice may influence the equilibrium
  structure: initial pore spacing, initial pore radius and initial distance
  between layers. Here we show the results of a sensitivity analysis performed
  on the three parameters. To reduce the size of the sensitivity analysis, we
  only tested systems built with 5 monomers per layer with layers stacked in 
  the parallel displaced configuration. We equilibrated all systems according
  to the dry equilibration procedure.

  %\begin{enumerate}

  %	  \item \underline{Initial pore spacing}
  \subsection{Initial pore spacing}\label{section:initial_pore_spacing}

	  We tested five different initial pore spacings, defined as the
	  distance between the central axis of each pore with all others
	  (Figure~\ref{fig:p2p_diagram}). To reduce the number of variables, we held the
	  pore radius constant, at 6 \AA~and the distance between layers at 3.7 \AA~since
	  those were the values used in our optimal system in the main text. We
	  prioritized ensuring that resulting configurations maintain the expected
	  hexagonal symmetry. If initial pore spacing is too small, we observe repulsion
	  between columns which disrupts equilibration. If the pore spacing is too large
	  the pores squeeze together, but in a distorted hexagonal array because the xy
	  initial translation of the pores is somewhat erratic.

@@

		\item 55 \AA~: We tested a pore spacing of 55 \AA~which is at
		a distance where monomers in each pore no longer intersect those of adjacent pores. Once
		the restraining potential reaches 56 kJ mol$^{-1}$ nm$^{-2}$, the pore spacing
		changes erratically until it begins to settle when the force constants are
		below 3 kJ mol$^{-1}$ nm$^{-2}$ (Figure~\ref{fig:p2p_55}). We recommend
		avoiding a system such as this where vacuum gaps between pore columns are
		introduced unnecessarily. %The system equilibrates to a pore spacing of 4.27
%		$\pm$ 0.18 nm. The 1-2 pore spacing is the outlier in this case with a pore
 %               spacing of 3.99 \AA~ which distorts the hexagonal array.

@@

	  \begin{figure}[!htb]
		\centering
		\includegraphics[width=0.4\textwidth]{p2p_diagram.png}
		\caption{In a perfect hexagonal, there are 5 distances that should be exactly
		equal. As pictured in this diagram the distance from 1-2, 1-3, 1-4, 2-4, and 3-4
		are the same.}~\label{fig:p2p_diagram}
	  \end{figure}

	  %BJC: Could put vertical lines to indicate where position restraint values are reduced
  	  \begin{figure}[!htb]
		\centering
		\begin{subfigure}{0.3\textwidth}
			\includegraphics[width=\textwidth]{p2p_39.png}\quad
			\vspace{-1.25em}
			\caption{39 \AA~initial pore spacing}~\label{fig:p2p_39}
		\end{subfigure}
		\begin{subfigure}{0.3\textwidth}
			\includegraphics[width=\textwidth]{p2p_41.png}\quad
			\vspace{-1.25em}
			\caption{41 \AA~initial pore spacing}~\label{fig:p2p_41}
		\end{subfigure}
		\begin{subfigure}{0.3\textwidth}
			\includegraphics[width=\textwidth]{p2p_45.png}
			\vspace{-1.25em}
			\caption{45 \AA~initial pore spacing}~\label{fig:p2p_45}
		\end{subfigure}
		\begin{subfigure}{0.3\textwidth}
			\includegraphics[width=\textwidth]{p2p_50.png}\quad	
			\vspace{-1.25em}
			\caption{50 \AA~initial pore spacing}~\label{fig:p2p_50}
		\end{subfigure}
		\begin{subfigure}{0.3\textwidth}
			\includegraphics[width=\textwidth]{p2p_55.png}
			\vspace{-1.25em}
			\caption{55 \AA~initial pore spacing}~\label{fig:p2p_55}
		\end{subfigure}

		\caption{The pore spacing during the restrained portion of the
			dry equilibration procedure is shown. Every 50 ps (0.05 ns) position restraints
			are reduced according to the sequence: 1000000, 3162, 56, 8, 3, 2, 1, 0 kJ
			mol$^{-1}$ nm$^{-2}$. When the initial pore spacing is chosen below the
			experimental value (a) or at the experimental value (b), there is an abrupt
			change in pore spacing when the position restraints are reduced to 56 kJ
			mol$^{-1}$ nm$^{-2}$. When pores are started 45 \AA~apart (c) the pore spacing
			remains relatively stable. When pores are spaced 50 \AA~apart (d), the pore
			spacing decreases nearly linearly once the restraints are reduced to 56 kJ
			mol$^{-1}$ nm$^{-2}$. When pores are spaced 55 \AA~apart (e), so that monomers
			do not intersect with adjacent pores, and position restraints are reduced to 56
			kJ mol$^{-1}$ nm$^{-2}$, the pore spacing changes erratically before stabilizing
			when force constants are reduced below 3 kJ mol$^{-1}$ nm$^{-2}$.} 
		\label{fig:p2p}
	  \end{figure}

%	  \item \underline{Initial pore radius}
	  \subsection{Initial pore radius}\label{section:initial_pore_radius}

	  We tested 3 different pore radii, defined as shown in
	  Figure~\ref{fig:pore_radius_illustration}. For each system, we held the initial
	  pore spacing constant at 45 \AA~and the initial distance between layers
	  constant at 3.7 \AA. Equilibrated values of pore radii are presented in
	  Table~\ref{table:radii}. The pore radius in each simulation frame is calculated
	  as the average distance of all carbonyl carbons (see
	  Figure~\ref{fig:pore_radius_illustration}) from their associated pore center.
	  Statistics for the pore radii reported were generated from the time series
	  representing the average pore radius at each frame. We detected equilibration
	  using \texttt{pymbar.timeseries.detectEquilibration}. We calculated the average
	  and standard deviation of pore radii using only data points collected after
	  equilibration was detected.

@@

		\item 8 \AA~: The largest pore radius that can achieved before
		energy minimization becomes problematic is 8 \AA. One should use caution with
		such a structure because of the relatively large vacuum space that is created
		in the pore region of the initial configuration. After being run through the
		full dry equilibration procedure, we see a combination of cylindrical and
		slit-like pores (Figure~\ref{fig:slits}). Measuring the pore radius of this
		system does not have a concrete meaning since slits do not have a single
		radius, but its calculated value is reported in Table ~\ref{table:radii}. We
		are wary of such a non-symmetric structure and choose not to use a pore radius
		of 8 \AA~for our starting configurations.

@@

	  \begin{figure}[!htb]
	  \centering
          \includegraphics[width=\textwidth]{slit_pores.png}
	  \caption{A system that was built with an initial pore radius of 8
		  \AA~equilibrates to a structure that exhibits both cylindrical and slit-like
		  pores. As pictured here, sodium ions are colored blue, carbon atoms in the
		  aromatic ring of the head group are colored orange and all else is colored
		  cyan.}~\label{fig:slits}
	  \end{figure}

%	  \item \underline{Initial distance between layers}
	  \subsection{Initial distance between layers}\label{section:initial_dbwl}

	  We tested 3 different initial layer spacings, defined as the distance
	  between the planes of aromatic rings in each layer. Systems built with layers
	  stacked 3.7 \AA~and 5 \AA~apart are discussed extensively in the main text. We
	  also tested a system built with layers stacked 10 \AA~ apart.

@@

	  \begin{figure}[!htb]
		\centering
		\includegraphics[width=0.5\textwidth]{dbwl_10.png}
		\caption{When layers are initially stacked 10 \AA~apart and the system
                is equilibrated using the dry equilibration procedure, large vacuum gaps
		form as the monomers attempt to fill space.}\label{fig:dbwl_10} 
	  \end{figure}

%  \end{enumerate}

%  \noindent
%  \begingroup
%        \fontsize{14pt}{14pt}\selectfont
%	\textbf{Calculation of pore-to-pore spacing statistics}
%  \endgroup

  \begin{figure}[!htb]
	\centering
	\includegraphics[width=0.25\textwidth]{p2p_diagram.png}
	\caption{There are five pore-to-pore distances that are equal in a perfect
		hexagon. If each number in the diagram represents the center of a
		pore in a hexagonal unit cell, then the distance from 1 to 2, 2 to 4, 
		4 to 3, 3 to 1 and 1 to 4 should be equal. Only 4 of the distances
		are independent. For example, the distance from 1 to 4 is defined by
		the location of all other pore centers.}
	\label{fig:p2p_diagram}
  \end{figure}

@@

  \begin{figure}[!htb]

  \begin{subfigure}{\linewidth}
        \centering
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{layered_nematic_order.png}
                \caption{Sandwiched, Ordered}~\label{fig:sandwich_nematic}
        \end{subfigure}%
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{offset_nematic_order.png}
                \caption{Parallel Displaced, Ordered}~\label{fig:offset_nematic}
        \end{subfigure}
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_sandwich_nematic_order.png}
                \caption{Sandwiched, Disordered}~\label{fig:disorder_sandwich_nematic}
        \end{subfigure}%
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_offset_nematic_order.png}
                \caption{Parallel Displaced, Disordered}~\label{fig:disorder_offset_nematic}
        \end{subfigure}

  \end{subfigure}
  \caption{The distribution of angles between nematic director vector (see
	  Figure~\ref{fig:director}) and the z-axis averaged over the equilibrated portion
	  of each trajectory}~\label{fig:nematic_distribution}
  \end{figure}

  \section{Measurement of intensities of experimental and simulated XRD patterns}\label{section:xrd_intensities}

  We developed standardized methods of measuring the intensity of reflections of interest
  in the experimental and simulated X-ray diffraction patterns presented in this work. Here,
  we graphically illustrate the methods used for measuring these intensities.

  All diffraction patterns are normalized according to R-alkanes and the intensities
  of all reflections are reported relative to the intensity of R-alkanes. The region
  used to calculate the average intensity of R-alkanes is shown in 
  Figure~\ref{fig:ralkanes}. The top region is excluded because, in the simulated
  patterns, R-$\pi$ intersects with R-alkanes.

  R-$\pi$ and R-double are reported as the maximum values of the peaks of the 
  $q_z$ cross-section of the experimental pattern at $q_r=0$ (Figure~\ref{fig:rpi_rdouble}).
  In the case of the experimental pattern, the peak heights are not symmetrical,
  so we report the average of the two heights.

>>>>

  We parameterized monomers according to the following procedure:
  \begin{enumerate}
	\item \textit{Create monomer structure file with connectivity} : We
	drew atomistic structures using MarvinSketch
	17.13~\cite{chemaxon_marvinsketch_2017} with all hydrogen atoms drawn out
	explicitly. We optimized the 3D geometry of the structure using the 'Clean in
	3D' function of MarvinSketch.  We saved the structure as a .mol file, then
	converted it to .pdb format using Open Babel 2.4.1
	\cite{oboyle_open_2011,noauthor_open_nodate}. 
	\item \textit{Assign GAFF atomtypes using \texttt{antechamber}} : Using
	the .pdb structure file as input, we ran
	\texttt{antechamber}~\cite{wang_automatic_2006} using the AM1-BCC charge model.
	The net charge on the monomer is input as -1 since the sodium ion is kept as 
	a separate residue. We use \texttt{LEaP}~\cite{case_ambertools16_2016} and the
	output of \texttt{antechamber} to create Amber topology files. A detailed
	tutorial can be accessed elsewhere \cite{walker_antechamber_nodate}.
	\item \textit{Create GROMACS topologies from Amber output} : The output
	of \texttt{LEaP} is a .inpcrd and a .prmtop file which are Amber topology
	files. Using acpype.py \cite{sousa_da_silva_acpype_2012}, we converted the
	\texttt{LEaP} output into GROMACS .gro and .top files. 
	\item \textit{Perform a simulated annealing procedure on the monomer} :
	We created a cubic box around the monomer using the GROMACS command \texttt{gmx
	editconf}. The monomer was centered in the box with edges of the
	box spaced at least 3 nm from the monomer on all sides. We ran an energy minimization
	on the system with the steepest descent algorithm. Next we performed an NVT
	simulated annealing procedure. We linearly decreased the temperature of the
	system from 1000 K to 50 K over the course of 10 ns. We randomly chose a monomer
	configuration from the last 10 \% of the trajectory. 
	\item \textit{Reassign charges with \texttt{molcharge}}: With the monomer
	configuration taken from the annealed trajectory, we reassigned charges using
	\texttt{molcharge} with the am1bccsym method in order to ensure charges 
	are symmetric. This condition is not guaranteed with \texttt{antechamber}. 
	The charges in the GROMACS topology file (.top) were replaced with the 
	new charges calculated by \texttt{molcharge}. 
	\item \textit{Anneal again to get final structure} : We repeated the
	same simulated annealing procedure using the monomer topology with
	\texttt{molcharge} charges. A random monomer configuration was pulled from the
	last 10 \% of the trajectory and was used to build all assemblies reported
	(Figure~\ref{fig:monomer}).
  \end{enumerate} 
  \clearpage
  \begin{figure}[!htb]
	\centering
        \includegraphics[width=0.9\textwidth]{monomer.png}
	\caption{Atomistic representation of the monomer Na-GA3C11. White atoms
		represent hydrogen, cyan atoms represent carbon, red atoms represent oxygen and
		the blue atom is sodium.}\label{fig:monomer}
  \end{figure}

@@

  \section{Initial Configuration Dependence}\label{section:initial_config_dependence}

  We addressed any major dependence on initial configuration in the main text.
  There we showed that systems are stable when made with 4, 5, 6, 7, and 8
  columns per pore. We also showed that systems are stable when monomer head
  groups are oriented in the parallel displaced and sandwiched configurations
  (see Figure~\ref{fig:stacking}). Our model is most consistent with experiment
  when built with 5 columns per pore, with monomers stacked 3.7 \AA~apart in 
  the parallel displaced configuration.

  There are three other parameters whose choice may influence the equilibrium
  structure: initial pore spacing, initial pore radius and initial stacking distance
  between monomers. Here we show the results of a sensitivity analysis performed
  on the three parameters. To reduce the size of the sensitivity analysis, we
  only tested systems built with 5 columns per pore with monomers stacked in 
  the parallel displaced configuration. We equilibrated all systems according
  to the dry equilibration procedure.

  \subsection{Initial pore spacing}\label{section:initial_pore_spacing}

  We tested five different initial pore spacings, defined as the
  distance between the central axis of each pore with all others
  (Figure~\ref{fig:p2p_diagram}). To reduce the number of variables, we held the
  pore radius constant, at 5 \AA~and the distance between layers at 3.7 \AA~since
  those were the values used in our optimal system in the main text. We
  prioritized ensuring that resulting configurations maintain the expected
  hexagonal symmetry. If initial pore spacing is too small, we observe repulsion
  between columns which disrupts equilibration. If the pore spacing is too large
  the pores squeeze together, but in a distorted hexagonal array because the $xy$
  initial translation of the pores is somewhat erratic.

@@

	\item 55 \AA~: We tested a pore spacing of 55 \AA~which is at
	a distance where monomers in each pore no longer intersect those of adjacent pores. Once
	the restraining potential reaches 56 kJ mol$^{-1}$ nm$^{-2}$, the pore spacing
	changes erratically until it begins to settle when the force constants are
	below 3 kJ mol$^{-1}$ nm$^{-2}$ (Figure~\ref{fig:p2p_55}). We recommend
	avoiding a system such as this where vacuum gaps between pore columns are
	introduced unnecessarily. 
	%The system equilibrates to a pore spacing of 4.27
    %$\pm$ 0.18 nm. The 1-2 pore spacing is the outlier in this case with a pore
    %spacing of 3.99 \AA~ which distorts the hexagonal array.

@@

	% BJC: Could put vertical lines to indicate where position restraint values are reduced
  	\begin{figure}[!htb]
	\centering
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\textwidth]{p2p_39.png}\quad
		\vspace{-1.25em}
		\caption{39 \AA~initial pore spacing}~\label{fig:p2p_39}
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\textwidth]{p2p_41.png}\quad
		\vspace{-1.25em}
		\caption{41 \AA~initial pore spacing}~\label{fig:p2p_41}
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\textwidth]{p2p_45.png}
		\vspace{-1.25em}
		\caption{45 \AA~initial pore spacing}~\label{fig:p2p_45}
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\textwidth]{p2p_50.png}\quad	
		\vspace{-1.25em}
		\caption{50 \AA~initial pore spacing}~\label{fig:p2p_50}
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\textwidth]{p2p_55.png}
		\vspace{-1.25em}
		\caption{55 \AA~initial pore spacing}~\label{fig:p2p_55}
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
	\includegraphics[width=\textwidth]{p2p_diagram.png}
	\caption{}~\label{fig:p2p_diagram}
%		\caption{In a perfect hexagonal, there are 5 distances that should be exactly
%		equal. As pictured in this diagram the distance from 1-2, 1-3, 1-4, 2-4, and 3-4
%		are the same.}~\label{fig:p2p_diagram}
  	\end{subfigure} 
	\caption{The pore spacing during the restrained portion of the
	dry equilibration procedure is shown. Every 50 ps (0.05 ns) we reduce the 
	force constant on the position restraints according to the sequence: 1000000,
	3162, 56, 8, 3, 2, 1, 0 kJ mol$^{-1}$ nm$^{-2}$. When we chose an initial pore
	spacing below experimental value (a) or at the experimental value (b), there
	is an abrupt change in pore spacing when the position restraints are reduced to 56 kJ
	mol$^{-1}$ nm$^{-2}$. When we initially position pores 45 \AA~apart (c) the pore spacing
	remains relatively stable. When we initially place pores 50 \AA~apart (d), the pore
	spacing decreases nearly linearly once the restraints are reduced to 56 kJ
	mol$^{-1}$ nm$^{-2}$. When we initially place pores 55 \AA~apart (e), so that monomers
	do not intersect with adjacent pores, and position restraints are reduced to 56
	kJ mol$^{-1}$ nm$^{-2}$, the pore spacing changes erratically before stabilizing
	when force constants are reduced below 3 kJ mol$^{-1}$ nm$^{-2}$. (f)
	In a perfect hexagonal, there are 5 distances that should be exactly
	equal. As pictured in this diagram the distance from 1-2, 1-3, 1-4, 2-4, and 3-4
	are the same. All plots in (a)--(e) are the average of these distances.} 
	\label{fig:p2p}
  \end{figure}

  \subsection{Initial pore radius}\label{section:initial_pore_radius}

  We tested 3 different pore radii, defined as shown in
  Figure~\ref{fig:pore_radius_illustration}. For each system, we held the initial
  pore spacing constant at 45 \AA~and the initial stacking distance between monomers
  constant at 3.7 \AA. Equilibrated values of pore radii are presented in
  Table~\ref{table:radii}. The pore radius in each simulation frame is calculated
  as the average distance of all carbonyl carbons (see
  Figure~\ref{fig:pore_radius_illustration}) from their associated pore center.
  We generated statistics for the pore radii reported from the time series
  representing the average pore radius at each frame. We detected equilibration
  as described in Section~\ref{M-method:equil_time} of the main text. We 
  calculated the average and standard deviation of pore radii using only data 
  points collected after equilibration was detected.

@@

  \item 8 \AA~: The largest pore radius that can achieved before
  energy minimization becomes problematic is 8 \AA. One should use caution with
  such a structure because of the relatively large vacuum space that is created
  in the pore region of the initial configuration. After being run through the
  full dry equilibration procedure, we see a combination of cylindrical and
  slit-like pores (Figure~\ref{fig:slits}). Measuring the pore radius of this
  system does not have a concrete meaning since slits do not have a single
  radius, but its calculated value is still reported in Table ~\ref{table:radii}. We
  are wary of such a non-symmetric structure and choose not to use a pore radius
  of 8 \AA~for our starting configurations.

@@

  \begin{figure}[!htb]
  \centering
  \begin{minipage}{0.475\textwidth}
  \centering
  \includegraphics[width=0.8\textwidth]{pore_radius_illustration.png}
  \caption{When creating an initial configuration, we define the pore radius,
  $\mathbf{R}$, based on the distance of the carbonyl carbon from the pore's
  central axis.}~\label{fig:pore_radius_illustration}
  \end{minipage}\qquad
  \begin{minipage}{0.475\textwidth}
  \centering
  \vspace{4em}
  \includegraphics[width=\textwidth]{slit_pores.png}
  \vspace{1em}
  \caption{A system that was built with an initial pore radius of 8~\AA~equilibrates
  to a structure that exhibits both cylindrical and slit-like pores. As pictured 
  here, sodium ions are colored blue, carbon atoms in the aromatic ring of the head 
  group are colored orange and all else is colored cyan.}~\label{fig:slits}
  \end{minipage}
  \end{figure}

  \subsection{Initial stacking distance between monomers}\label{section:initial_dbwl}

  We tested 3 different initial monomer spacings, defined as the $z$-distance
  between the planes of aromatic rings. Systems built with layers
  stacked 3.7 \AA~and 5 \AA~apart are discussed extensively in the main text. We
  also tested a system built with layers stacked 10 \AA~ apart.

@@

  \begin{figure}[!htb]
	\centering
	\includegraphics[width=0.3\textwidth]{dbwl_10.png}
	\caption{When layers are initially stacked 10 \AA~apart and the system
                is equilibrated using the dry equilibration procedure, large vacuum gaps
	form as the monomers attempt to fill space.}\label{fig:dbwl_10} 
  \end{figure}

  \clearpage
  \section{Placement of monomer head groups}

  The monomer head groups consist of a substituted benzene ring which may participate in
  $\pi-\pi$ stacking interactions. There are three $\pi-\pi$ stacking modes known
  to occur in nature: sandwiched, parallel displaced and T-shaped
  (Figures~\ref{fig:sandwiched}--\ref{fig:tshaped})~\cite{sinnokrot_estimates_2002}.
  We ruled out the T-shaped configuration because its vertical stacking distance is
  inconsistent with experimental spectroscopy. Additionally, steric interactions
  between the bulky alkane tails make this motif unfavorable. We chose to explore
  systems in both the sandwiched and parallel displaced configurations since their
  stacking distances are both consistent with experimental diffraction patterns.

  \begin{figure}[!htb]
        \centering
        \begin{subfigure}[b]{0.32\textwidth}
                \centering
                \includegraphics[width=\textwidth]{sandwiched.png}
                \caption{}\label{fig:sandwiched}
        \end{subfigure}
        \begin{subfigure}[b]{0.32\textwidth}
                \centering
                \includegraphics[width=\textwidth]{PD.png}
                \caption{}\label{fig:pd}
        \end{subfigure}
        \begin{subfigure}[b]{0.32\textwidth}
                \centering
                \includegraphics[width=\textwidth]{Tshaped.png}
                \caption{}\label{fig:tshaped}
        \end{subfigure}
        \vskip\baselineskip
        \begin{subfigure}[b]{0.475\textwidth}
                \centering
                \includegraphics[width=\textwidth]{sandwichedlayers.png}
                \caption{}\label{fig:sandwichedlayers}
        \end{subfigure}
        \begin{subfigure}[b]{0.475\textwidth}
                \centering
                \includegraphics[width=\textwidth]{offsetlayers.png}
                \caption{}\label{fig:offsetlayers}
        \end{subfigure}
        \caption{(a) Sandwiched benzene dimers stack 3.8 \angstrom~apart. (b) Parallel-Displaced benzene dimers stack
        3.4 \angstrom~vertically and 1.6 \angstrom~horizontally apart. (c) T-shaped benzene dimers stack 5.0 \angstrom~apart.
        (d) Monomers stacked in the sandwiched configuration (e) Monomers stacked in the parallel-displaced
        configuration.}\label{fig:stacking}
  \end{figure}

  We chose to build all systems using 20 monomers per column. We made this choice
  by considering a number of factors that directly affect the quality of our results. 
  Clearly, choosing a small number of layers will cause the system to suffer from 
  finite size effects. Even in our 20 monomer-per-column system, axial correlations
  between stacked monomers persist throughout the entire correlation function (see
  Figure ~\ref{M-fig:correlation}). The correlation function of the system which we
  built with 40 monomers per column is very similar to that created with 20 monomers
  per columns (continues to oscillate for its full length
  (Figure~\ref{fig:tall_correlation_full}). If we continue adding monomers, the 
  correlation function will eventually fully decay but it is not worth the 
  computational expense. Regardless, the correlation lengths calculated from 20
  monomer per column simulations are in sufficient agreement with experiment.

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{z_correlation_overlay.png}
  \caption{}\label{fig:z_correlation_overlay}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{tall_correlation_full.png}
  \caption{}\label{fig:tall_correlation_full}
  \end{subfigure}
  \caption{(a) The correlation functions generated from ordered basin sandwiched 
  configurations are nearly the same when the $z$-dimension of the system is doubled.
  (b) Oscillations persist throughout the length of the correlation function for the
  sandwiched system in the ordered basin built with 40 monomers per column. The full
  system is ca. 17 nm tall so the full correlation function covers half this length
  since periodicity forces the remaining length to be its mirror image.}\label{fig:correlation_functions}
  \end{figure}

  We may be able to simulate systems using less than 20 monomers per column, but chose
  to use 20 since it gives fairly good resolution when simulating X-ray diffraction 
  patterns. The size of the Fourier space bin in each dimension is determined by 
  $\dfrac{2\pi}{L}$ where L is the length of a given box vector. The $z$-box vector 
  for 20 monomer-per-column systems is $\approx$ 85 \AA which means we will see 
  a reciprocal space resolution of 0.074 \AA$^{-1}$ in the z direction. Further decreasing
  the resolution would increase the noise in the simulated patterns and make it more
  difficult to measure the intensity of reflections of interest.

@@

  \clearpage

  \begin{figure}[!htb]
  \begin{subfigure}{\linewidth}
        \centering
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{layered_nematic_order.png}
                \caption{Sandwiched, Ordered}~\label{fig:sandwich_nematic}
        \end{subfigure}%
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{offset_nematic_order.png}
                \caption{Parallel Displaced, Ordered}~\label{fig:offset_nematic}
        \end{subfigure}
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_sandwich_nematic_order.png}
                \caption{Sandwiched, Disordered}~\label{fig:disorder_sandwich_nematic}
        \end{subfigure}%
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_offset_nematic_order.png}
                \caption{Parallel Displaced, Disordered}~\label{fig:disorder_offset_nematic}
        \end{subfigure}
  \end{subfigure}
  \caption{The distribution of angles between nematic director vector (see
	  Figure~\ref{fig:director}) and the z-axis averaged over the equilibrated portion
	  of each trajectory}~\label{fig:nematic_distribution}
  \end{figure}
